---
title: "Hands on Practice"
author: "R Trainer: Binod Jung Bogati (linkedin.com/in/bjungbogati)"
copyright: "© 2025. Unauthorized distribution or reuse prohibited."
format: html
editor: visual
---

```{r}
library(tidyverse)

load(url("https://bit.ly/4eZIXte"))
```

## Your Turn 01

Plot histogram of `Height` using `vs` Dataset.

Add `binwidth` = 5.

```{r}
ggplot(...., aes(x = ....)) +
  geom_histogram(binwidth = ....)
```

Again plot the histogram of `Temp` using vs dataset:

Add bindwidth = 0.1 and fill = "skyblue"

```{r}





```

## Your Turn 02

Draw a bar plot to find out the number of subjects in each ARMCD. Use `dm` dataset.

Use following properties:

-   Geometry: fill = "steelblue"
-   Title : "Subject Distribution by Treatment Arm",
-   X axis : "Treatment Arm Code (ARMCD)",
-   Y axis : "Number of Subjects",
-   Caption: "Source: DM dataset"
-   Add theme_minimal()

```{r}
ggplot(dm, aes(x = ...))+ 
  ...

```

## Your Turn 03

Verify the count of the subjects in each `ARMCD` by using summarise and n() functions learned on Day 6.

```{r}
count_ex3 <- dm |>
  group_by(...) |>
  summarize(...)

# alternatively we can use count() function
count_ex3 <- dm |>
  count(...)
```

## Your Turn 04

Draw a line chart using the dataset of `vs_facet` to show `VISIT` in x-axis, `mean` in y-axis and group and colored by `ARMCD` and faceted by `VSTESTCD`.

-   Use `theme_classic()`
-   Add scales = "free"

Use `theme_classic()`

> facet means displaying same plot with a different groups 

```{r}

ggplot(...., aes(x = , y = , color = , group = )) +
  geom_.....() +
  geom_point() +
  facet_wrap(........, scales = "free") +
  theme_classic()
```
