---
title: "Hands on Practice"
author: "R Trainer: Binod Jung Bogati (linkedin.com/in/bjungbogati)"
copyright: "© 2025. Unauthorized distribution or reuse prohibited."
format: html
editor: visual
---

```{r}
library(tidyverse)

load(url("https://bit.ly/4eZIXte"))
```

## Your Turn 01

Plot histogram of `Height` using vs Dataset

```{r}
ggplot(vs, aes(x = ....)) +
  geom_histogram(binwidth = 5)
```

Again plot the histogram of `Temp` using vs dataset:

```{r}
ggplot(vs, aes(x = ...)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue") 
```

## Your Turn 02

Draw a bar plot to find out the number of subjects in each ARMCD. Use `dm` dataset.

```{r}
ggplot(dm, aes(x = ...))+ 
  ...

```

## Your Turn 03

Verify the count of the subjects in each `ARMCD` by using summarise and n() functions learned on Day 6.

```{r}
count_ex3 <- dm |>
  group_by(...) |>
  summarize(...)

# alternatively we can use count() function
count_ex3 <- dm |>
  count(...)
```

## Your Turn 04

Draw a line chart using the dataset of `vs_facet` to show `VISIT` in x-axis, `mean` in y-axis and group and colored by `ARMCD` and faceted by `VSTESTCD`.

Use `theme_classic()`

> facet means displaying same plot with a different groups 

```{r}

ggplot(...., aes(x = , y = , color = , group = )) +
  geom_.....() +
  geom_point() +
  facet_wrap(........, scales = "free") +
  theme_classic()
```
